This is an example of article title

### 2019-05-25 paper/paper.tex
vectors transpose

\section{SIMD sorting methods}
\label{sec:methods}
%
ChipSort offers three high-level functions that can be used in applications as replacements for other sorting functions such as \verb sort  in the Julia standard library. Additional work is required from the user selecting among the three functions and also setting some numeric parameters. The functions are intended to be used with arrays of different sizes, and \verb chipsort_large  offers a good overview of all the techniques implemented in the module. Some of these techniques might also be useful in other applications instead of sorting, one clear example being the in-place matrix transpose. ChipSort does not currently support registers with separate key and payload, what might allow significant optimizations. Only the sorting of simple data types such as 32 and 64-bit integers and floating-point numbers were studied.

@@

A sorting network can be represented by a Knuth graph such as Figure~\ref{fig:sorting-network}. The graph represents a network that takes an array of four values $(A_1, A_2, A_3, A_4)$ and outputs the sorted sequence $(S_1, S_2, S_3, S_4)$. Each vertical line in the graph represents a comparator.
\begin{figure}[htb]
\centerline{\includegraphics[width=0.9\linewidth]{fig/sorting-network-4.pdf}}
\caption{A sorting network for 4 elements. The dashed lines represent sections where operations are independent of each other.}
\label{fig:sorting-network}
\end{figure}

@@

\begin{figure}[htb]
\begin{lstlisting}[language = Julia]
input_0_1 = Base.getindex(input, 1)
input_0_2 = Base.getindex(input, 2)
input_0_3 = Base.getindex(input, 3)
input_0_4 = Base.getindex(input, 4)
input_1_1 = ChipSort.min(input_0_1, input_0_2)
input_1_2 = ChipSort.max(input_0_1, input_0_2)
input_1_3 = ChipSort.min(input_0_3, input_0_4)
input_1_4 = ChipSort.max(input_0_3, input_0_4)
input_2_1 = ChipSort.min(input_1_1, input_1_3)
input_2_3 = ChipSort.max(input_1_1, input_1_3)
input_2_2 = ChipSort.min(input_1_2, input_1_4)
input_2_4 = ChipSort.max(input_1_2, input_1_4)
input_3_1 = input_2_1
input_3_4 = input_2_4
input_3_2 = ChipSort.min(input_2_2, input_2_3)
input_3_3 = ChipSort.max(input_2_2, input_2_3)
return (input_3_1, input_3_2, input_3_3, input_3_4)
\end{lstlisting}
\caption{Code generated by {\tt sort\_net} to implement a 4-element sorting network.}
\label{fig:sort-net-listing}
\end{figure}

@@

\subsection{SIMD}

When applied to a group of vectors, the result from the sorting network contains sorted sequences in the lanes of the vectors. To obtain sorted vectors we need to transpose these vectors.

>>>>

\section{SIMD sorting techniques}
\label{sec:methods}
%
ChipSort offers three high-level functions that can be used in applications as replacements for other sorting functions such as \verb sort  in the Julia standard library. Additional work is required from the user selecting among the three functions and also setting some numeric parameters. The functions are intended to be used with arrays of different sizes, and \verb chipsort_large  offers a good overview of all the techniques implemented in the module. Some of these techniques might also be useful in other applications instead of sorting, one clear example being the in-place matrix transpose. ChipSort does not currently support registers with separate key and payload, what might allow significant optimizations. Only the sorting of simple data types such as 32 and 64-bit integers and floating-point numbers were studied.

@@

A sorting network can be represented by a Knuth graph such as Figure~\ref{fig:sorting-network}. The graph represents a network that takes an array of four values $(A_1, A_2, A_3, A_4)$ and outputs the sorted sequence $(S_1, S_2, S_3, S_4)$. Each vertical line in the graph represents a comparator.
\begin{figure}[htb]
\centerline{\includegraphics[width=0.6\linewidth]{fig/sorting-network-4.pdf}}
\caption{A sorting network for 4 elements. The dashed lines delimit each step where operations are independent of each other.}
\label{fig:sorting-network}
\end{figure}

@@

\begin{figure}[htb]
\begin{lstlisting}[language = Julia]
4 => (((1,2), (3,4)), ((1,3), (2,4)), ((2,3),))
\end{lstlisting}
\begin{lstlisting}[language = Julia]
input_0_1 = input[1]
input_0_2 = input[2]
input_0_3 = input[3]
input_0_4 = input[4]
input_1_1 = min(input_0_1, input_0_2)
input_1_2 = max(input_0_1, input_0_2)
input_1_3 = min(input_0_3, input_0_4)
input_1_4 = max(input_0_3, input_0_4)
input_2_1 = min(input_1_1, input_1_3)
input_2_3 = max(input_1_1, input_1_3)
input_2_2 = min(input_1_2, input_1_4)
input_2_4 = max(input_1_2, input_1_4)
input_3_1 = input_2_1
input_3_4 = input_2_4
input_3_2 = min(input_2_2, input_2_3)
input_3_3 = max(input_2_2, input_2_3)
return (input_3_1, input_3_2, input_3_3, input_3_4)
\end{lstlisting}
\caption{The parameters from a four-elements sorting network, and the corresponding code generated by {\tt sort\_net} to implement it.}
\label{fig:sort-net-listing}
\end{figure}

@@

\subsection{SIMD vectors transpose}

Consider a group of stacked SIMD vectors. Each column in this representation is termed a {\em lane}. The result from {\tt sort\_vecs} is that each lane contains a sorted sequence. To obtain sorted vectors we need to transpose these stacked vectors. This is done in ChipSort through {\tt transpose\_vecs}, which is a generated function that supports rectangular matrices, but with dimensions constrained to powers of 2. This function is heavily based on calls to {\tt SIMD.shufflevec}.

Figure~\ref{fig:transpose-vecs} displays an example of 64 random integers loaded into SIMD registers as 4 vectors. After {\tt sort\_vecs} each lane contains a sorted sequence. After {\tt transpose\_vecs} each vector (register) contains a sorted sequence.

\begin{figure}[htb]
\centerline{\includegraphics[width=0.99\linewidth]{fig/regs.pdf}}
\caption{Four vectors with four values, the result from a vectorized four-elements sorting network, and the transpose.}
\label{fig:transpose-vecs}
\end{figure}
